Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of solar-type stars

### 2017-08-21 paper/ms.tex
improve intro via DNS comments

\newcommand{\todo}[1]{{\color{blue}TODO: #1\\}}

@@

\author{David N. Spergel} % TODO: need ORCID
\affil{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane,
  Princeton, NJ 08544, USA}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave,
  New York, NY 10010, USA}
\author{Justin Myles} % TODO: need ORCID
\affil{Department of Astronomy, Yale University, 260 Whitney Ave, New Haven, CT 06511, USA}

@@

Wide binary stars are valuable tools for studying star and planet formation as well
as Galactic dynamics and chemical evolution.
In the context of studying the evolution of the Milky Way, they are useful for
two main reasons.
First, because wide binaries are weakly bound systems that may be tidally
disrupted by, e.g., field stars, molecular clouds, or the Galactic tidal field,
their statistics can be informative of the Galactic mass distribution.
For example, the separation distribution of halo binaries has been used to
constrain the mass of massive compact halo objects (\citealt{Yoo:2004aa}).
On the other hand, they are also prime targets to test the chemical tagging
hypothesis that stars from the same birthplace may be traced back using detailed
chemical abundance patterns as birth ``tag''s (\citealt{2002ARA&A..40..487F}).
A multiple-star system of any size that we believe are born together can serve
this purpose, with massive open clusters at one extreme.
Wide binaries are at the other extreme in that we may be less confident of the
coevality of individual systems but they are much more abundant than larger
systems like open clusters, rendering their statistics a meaningful indication
of whether the hypothesis works.

Wide binaries with similar or very different planetary architectures around the
component stars have been used to connect a host star's chemistry to its
planetary system.
A differential analysis may reveal chemical signatures related to planet
formation or accretion without any regard to Galactic chemical evolution as
long as the two stars are assumed to be born together with identical
composition.
For close-in giant planets, it has been established that the planet-metallicity
relation is primordial, and that the enhanced metallicity in giant
planet-hosting stars likely promotes the formation of giant planets by
increasing the availability of small particle condensates that form
planetesimals (\citealt{Fischer:2005aa}).
On the other hand, if host stars are polluted after their birth by rocky
planetary material with high refractory-to-volatile ratio, the convective
envelope of the stars may be enhanced in refractory elements (e.g., \elem{Fe})
compared to their initial state.
Thus, differences in planet formation or accretion in two otherwise identical
stars may imprint differences in chemical abundances that depend on the
condensation temperature (\Tcondens).

Several systems in which at least one star hosts at least one planet
have been studied so far with high resolution spectroscopy in this context
(\citealt{Teske:2013aa,Mack:2014aa,Liu:2014aa,Teske:2015aa,Saffe:2015aa,
  Ramirez:2015aa,Biazzo:2015aa,Mack:2016aa,Teske:2016aa,Teske:2016ab}).
The results and interpretations of these studies are varied:
while some systems appear to have undetectable differences
(see also \citealt{Desidera:2004aa,Gratton:2001aa}), other
studies have reported a \Tcondens-dependent difference in abundance
with higher-\Tcondens\ elements showing larger differences.
The increasing abundance difference with condensation temperature has been
attributed both to more yet-undiscovered rocky planets around the relatively
refractory-poor star and to late time accretion of refractory-rich planetary
material (\citealt{Ramirez:2015aa,Biazzo:2015aa}).
The observed differences are $\lesssim 0.1$~dex even in the most dramatic case,
and often at $\approx 0.05$~dex level making them challenging to detect even
with a careful analysis of high-resolution high signal-to-noise ratio spectra.
% maybe want to point out the former interpretation roots back to melendez+09
% and some doubt to that interpretation...

So far the strongest evidence for accretion of planetary material by a host
star comes from spectral analysis of polluted white dwarfs (WDs).
Because the gravitational settling times of elements heavier than \elem{He} in
atmosphere is much shorter than the WD cooling time, detection of metals likely
indicates the presence of a reservoir of dusty material around the WD.
Indeed, many of the polluted WDs host a dusty debris disk detected
in the infrared.
Some of the most dramatically polluted WDs show
surface abundances closely matched by rocky planetary material
with, e.g., bulk Earth composition, strongly arguing
that the disk formed from tidally disrupted minor planets
(\citealt{Zuckerman:2007aa,Klein:2010aa}).
Recently, transit signals from small bodies orbiting around a polluted WD
have been detected by \project{Kepler} adding further support to the picture
(\citealt{2015Natur.526..546V}).

Here, we report and discuss the discovery of a comoving pair of G stars,
HD~240430 and HD~240429, with unusual chemical abundance differences
that strongly suggest accretion of planetary material by the two stars.
Throughout the \documentname, we nickname the two stars \bizarreone\
(HD~240430) and \sunanalog\ (HD~240429).
We use the following convention for chemical abundances of stars: \elemH{X} is
the log ratio of number density of element \elem{X} to \elem{H} relative to the
solar value.
The absolute abundance of \elem{X} is $A(\elem{X}) = 12 + \log_{10}
(n_\elem{X}/n_\elem{H})$ where $n_\elem{X}$ is the number density of element
\elem{X}.
In \sectionname~\ref{sec:data} we present the astrometric and spectroscopic data
about the two stars relevant to the present discussion.
In \sectionname~\ref{sec:discussion} we discuss possible interpretations of
the abundance difference between the pair.
We summarize our discussions in \sectionname~\ref{sec:summary}.

@@

\todo{add microturb vel to tab 1}

>>>>

\newcommand{\todo}[1]{{\color{blue}TODO: #1\\}}
\newcommand*{\askjmb}[1]{{\bf #1}}

@@

\author{David N. Spergel}
\affil{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane,
  Princeton, NJ 08544, USA}
\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave,
  New York, NY 10010, USA}
\author{Justin Myles}
\affil{Department of Astronomy, Yale University, 260 Whitney Ave, New Haven, CT 06511, USA}

@@

Wide binary stars are valuable tools for studying star and planet formation as
well as Galactic dynamics and chemical evolution.
In the context of studying the evolution of the Milky Way, they are useful for
two main reasons.
First, because wide binaries are weakly bound systems that may be tidally
disrupted by, e.g., field stars, molecular clouds, or the Galactic tidal field,
their statistics can be informative of the Galactic mass distribution.
For example, the separation distribution of halo binaries has been used to
constrain the mass of massive compact halo objects
(\citealt{Yoo:2004aa,Quinn:2009,Allen:2014}).
They can also be used to test the ``chemical tagging'' hypothesis that stars
from the same birthplace may be traced back using detailed chemical abundance
patterns as birth tags (\citealt{2002ARA&A..40..487F}).
While any multiple-star system including massive open clusters can be used to
test the hypothesis, wide binaries have the advantage of being extremely
abundant, rendering their statistics a meaningful indication of whether the
hypothesis works.

Binary stars that form from the same birth cloud start with nearly identical
composition.
A differential analaysis of the chemical composition of binary stars can reveal
their history through the chemical signatures related to planet formation or
accretion regardless of the Galactic chemical evolution.
\askjmb{
For example, \citealt{Fischer:2005aa} showed that the higher metallicity in
stars hosting close-in giant planets compared to those without likely promotes
the formation of giant planets by increasing the availability of small particle
condensates that form planetesimals.
}
On the other hand, if host stars are polluted after their birth by rocky
planetary material with high refractory-to-volatile ratio, the convective
envelope of the stars may be enhanced in refractory elements (e.g., \elem{Fe})
compared to their initial state (e.g., \citealt{Pinsonneault:2001aa}).
Thus, differences in planet formation or accretion in two otherwise identical
stars may imprint differences in chemical abundances that depend on the
condensation temperature (\Tcondens).

High resolution spectroscopic studies of binary star systems hosting at least
one planet
(\citealt{Teske:2013aa,Mack:2014aa,Liu:2014aa,Teske:2015aa,Saffe:2015aa,
  Ramirez:2015aa,Biazzo:2015aa,Mack:2016aa,Teske:2016aa,Teske:2016ab})
have yielded varied results:
while some systems appear to have undetectable differences
(see also \citealt{Desidera:2004aa,Gratton:2001aa}), other
studies have reported a \Tcondens-dependent difference in abundance
with higher-\Tcondens\ elements showing larger differences.
A possible explanation for the increasing abundance difference with
condensation temperature is that more yet-undiscovered rocky planets formed
around the relatively refractory-poor star
(\citealt{Ramirez:2015aa,Biazzo:2015aa}).
Alternatively, late time accretion of refractory-rich planetary material to the
relatively refractory-rich star can also produce the trend.
The observed differences are $\lesssim 0.1$~dex even in the most dramatic case,
and often at $\approx 0.05$~dex level making them challenging to detect even
with a careful analysis of high-resolution, high signal-to-noise ratio spectra,
and differential analyses of two stars that are very similar in their
stellar parameters.

Spectral analysis of polluted white dwarfs (WDs) currently provides the
strongest evidence for accretion of planetary material by a host star
(\citealt{2003ApJ...596..477Z,2010ApJ...722..725Z,2014A&A...566A..34K};
see \citealt{2016NewAR..71....9F} for review).
Because the gravitational settling times of elements heavier than \elem{He} in
atmosphere is much shorter than the WD cooling time
(\citealt{1986ApJS...61..197P}), detection of metals likely indicates the
presence of a reservoir of dusty material around the WD.
Indeed, many of the polluted WDs host a dusty debris disk detected in the
infrared (\citealt{1987Natur.330..138Z,1990ApJ...357..216G,2005ApJ...635L.161R,
  2009ApJ...694..805F,2006ApJ...646..474K}).
Some of the most dramatically polluted WDs show
surface abundances closely matched by rocky planetary material
with, e.g., bulk Earth composition, strongly arguing
that the disk formed from tidally disrupted minor planets
(\citealt{Zuckerman:2007aa,Klein:2010aa}).
Recently, transit signals from small bodies orbiting around a polluted WD
have been detected by \project{Kepler} adding further support to the picture
(\citealt{2015Natur.526..546V}).

Here, we report and discuss the discovery of a comoving pair of G stars,
HD~240430 and HD~240429, with unusual chemical abundance differences that
strongly suggest accretion of rocky planetary material by one of the two stars,
HD~240430.
Throughout the \documentname, we nickname the two stars \bizarreone\
(HD~240430) and \sunanalog\ (HD~240429).
In Greek mythology, Kronos and Krios were sons of Uranos and Gaia.
Kronos notoriously devoured all of his children (except Zeus)
to prevent the prophecy that one day he will be overthrown by them.
We use the following convention for chemical abundances of stars: \elemH{X} is
the log ratio of number density of element \elem{X} to \elem{H} relative to the
solar value.
The absolute abundance of \elem{X} is $A(\elem{X}) = 12 + \log_{10}
(n_\elem{X}/n_\elem{H})$ where $n_\elem{X}$ is the number density of element
\elem{X}.
In \sectionname~\ref{sec:data} we present the astrometric and spectroscopic data
about the two stars relevant to the present discussion.
In \sectionname~\ref{sec:discussion} we discuss possible interpretations of
the abundance difference between the pair.
We summarize our discussions in \sectionname~\ref{sec:summary}.

@@
